---
title: "Creating web APIs with Plumber"
author: "Guilherme Lacerda & Hyebin Hong"
date: "11/04/2021"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# API
## What is an API and Web API?

* API stands for Application Programming and Web API is an API over the web
* functions as a mediator between clients and web services
* typically uses HTTP protocol for request messages and provides responses in the form of an XML or JSON file

## Some Web API features
* supports CRUD operations, which correspond to HTTP methods GET,POST,PUT and DELETE
* responses have headers, HTTP status code and message body
* supports multiple text formats such as XML, JSON, and etc. 

## Why is a Web API useful?
* can be called and consumed by any kind of application such as mobile app, desktop app, and etc. 
* not limited to use specific interface in order to communicate with users
* uses low bandwidth
* open source

# Plumber
## create APIs with R
‘Plumber’ uses annotations which are special comments(#*) and tags(e.g. @apiTitle) to turn any arbitrary R functions into API endpoints. 

Let’s walk through the various annotations that are important to create an API! 
 
## Plumber package functioning
```{r, message=F, comment=F}
#* @filter logger
function(req){
  cat(as.character(Sys.time()), "-",
    req$REQUEST_METHOD, req$PATH_INFO, "-",
    req$HTTP_USER_AGENT, "@", req$REMOTE_ADDR, "\n")
  plumber::forward()
}

#* @get /dyn/<name:str>/<age:[int]>/route
#* @serializer text
#* @parser none
#* @response 200 A sentence
function(name, age) {
  sprintf("%s is %i years old", name, age)
}
```

* Filters: used to handle (and eventually modify) incoming requests.
  + Typically, a request will pass through every filter before going to the endpoint (the @preempt tag opts specific filters out).
  + Common filter uses are collecting information about incoming requests and authentication.


* Parser: determine how Plumber parses (i.e., analyzes a string or text into logical syntactic components) the incoming request body.
  + By default Plumber parses the request body as JavaScript Object Notation (JSON).
  + Other parsers, including custom parsers, are identified using the @parser [parser name] tag.
  + All registered parsers can be viewed with registered_parsers().

* Endpoint: define the R code that is executed in response to incoming requests.
  + Correspond to HTTP methods and respond to incoming requests that match the defined method.
  + Methods: @get, @post, @put, @delete, ...

* Serializer: determine how to return results to the client.
  + Default serializes the R object into JSON, but custom serializers can be applied through @serializer [serializer name] tag.
  + E.g., #* @serializer csv

* Interface and programming

* Deployment: There are several ways to deploy APIs developed by Plumber. RStudio Connect is one of the easiest ways, allowing push-button publishing from the RStudio IDE. 